---
title: "Mental Models QPRESS"
author: "Erik Olsen"
date: "2/10/2020"
output: html_document
---

# Analysis of future scenarios for North Sea regional mental models
Scenarios were explored using press perturbation (QPRESS) for four mental models of the North Sea:
* Southern North Sea (Dutch stakeholders, 2018)
* Skagerrak (WGINOSE scientists, 2018)
* Kattegat (WKKEMSSP, 2019)
* Norwegian Trenc (Norwegian stakeholders, 2019)

## Libraries & functions
```{r libraries and functions, echo=FALSE, warning=FALSE}
require(data.table)
library("tcltk")
library("tcltk2")
library("XML")
library("Rcpp")
library("tidyverse")

devtools::install_github("SWotherspoon/QPress", build_vignettes = TRUE)
require(QPress)


#MM2Qpress
#Function for converting Mental modeller spreadsheets into Qpress links
#by Sean Lucey
MM2Qpress <- function(data){
  library(data.table)
  mental.sheet <- as.data.table(data)
  names(mental.sheet)[1] <- 'Box'
  n <- nrow(mental.sheet)
  box.names <- names(mental.sheet)[which(names(mental.sheet) != 'Box')]
  model <- c()
  for(i in 1:n){
    pos <- which(mental.sheet[i, .SD, .SDcol = box.names] > 0)
    if(length(pos) > 0){
      pos.interaction <- paste(mental.sheet[i, Box], '->', 
                               names(mental.sheet[i, pos + 1, with = F]), 
                               sep = '')
      model <- append(model, pos.interaction)
    }
    
    neg <- which(mental.sheet[i, .SD, .SDcol = box.names] < 0)
    if(length(neg) > 0){
      neg.interaction <- paste(mental.sheet[i, Box], '-*', 
                               names(mental.sheet[i, neg + 1, with = F]), 
                               sep = '')
      model <- append(model, neg.interaction)
    }
  }
  return(model)
}

```

## Import MM & create signed digraphs
When reading in files from excel format R may read spaces differently in rows and columms.  
To be safe:
1. either use node names with no blanks OR  
2. convert the names in the csv file to replace " " with "_"  
  
The below file has used option 2 to rename nodes
and copied to ensure rows and columns have the same names

```{r import MM, echo=FALSE, warning=FALSE}
#southern North Sea
SNS_dutch <- read.csv2("./data/2018_Dutch_Stakeholders/SouthNS_dutch.csv", dec=".")
SNS<-MM2Qpress(SNS_dutch)
SNS<-dput(SNS)
SNSModel<-parse.digraph(SNS)

#skagerrak
skagerrak<- read.csv2("./data/2018_WGINOSE/Skagerrak_190418.csv", dec=".")
sk<-MM2Qpress(skagerrak)
sk<-dput(sk)
skagerrakModel<-parse.digraph(sk)

#kattegat
kattegat<- read.csv2("./data/2019_Kattegat/Kattegat_WKKEMSSP_2019_final.csv", dec=".")
kattegat2<-MM2Qpress(kattegat)
kattegat2<-dput(kattegat2)
kattegatModel<-parse.digraph(kattegat2)

#Norwegian trench
#NorwTrench<- read.csv2("./data/2019_NorwTrench/NorwTrench_eng.csv", dec=".")
NorwTrench <- read_excel("./data/2019_NorwTrench/NorwTrench_eng_oct20.xlsx")
NorwTrench2<-MM2Qpress(NorwTrench)
NorwTrench2<-dput(NorwTrench2)
NorwTrenchModel<-parse.digraph(NorwTrench2)
```





## QPRESS analysis
#### Southern North Sea QPRESS model
```{r SNS QPRESS, echo=FALSE, warning=FALSE}
#doesnt work without this, self loops imposed
SNSel<-enforce.limitation(SNSModel)

#run the simulation
SNSsim<-system.simulate(10, SNSel)

#make the interactive plot
#impact.barplot(SNSsim)

#shiny interactive plot 
impact.barplot.shiny(SNSsim)

```

#### Skagerrak QPRESS model
```{r Skagerrak QPRESS, echo=FALSE, warning=FALSE}
skag_el<-enforce.limitation(skagerrakModel)

#run the simulation
skag_sim<-system.simulate(10, skag_el)

#make the interactive plot
impact.barplot.shiny(skag_sim)


```

#### Kattegat QPRESS model
```{r kattegat QPRESS, echo=FALSE, warning=FALSE}
katt_el<-enforce.limitation(kattegatModel)

#run the simulation
katt_sim<-system.simulate(10, katt_el)

#make the interactive plot
impact.barplot.shiny(katt_sim)


```


#### Norwegian trench QPRESS model
```{r Norw trench QPRESS, echo=FALSE, warning=FALSE}
Nor_el<-enforce.limitation(NorwTrenchModel)

#run the simulation
Nor_sim<-system.simulate(10, Nor_el)

#make the interactive plot
impact.barplot.shiny(Nor_sim)

```

### QPRESS scenarios results as .CSV files
Code borrowed from https://github.com/NOAA-EDAB/QNM/blob/master/looping_qpress_Rpath.R  

#### Southern North Sea 
```{r SNS QPRESS scenarios to CSV & barplot, echo=FALSE, warning=FALSE}

#mmod <- NorwTrenchModel
mmod <- SNSel

scenarios <- c("fisheries_increase", "fisheries_decrease")
qls <- c(1,-1)

A <- adjacency.matrix(mmod)
#A

## Function to generate the community matrix
s <- community.sampler(mmod)

for (j in 1:length(scenarios)){ 
  ## Function to check the validation condition
   # press <- press.validate(mmod, perturb=c(Fisheries=1), monitor=F)
  press <- press.validate(mmod, perturb=c(Fisheries=qls[j]), monitor=F)
  
  ## Function to define the perturbation scenario
  #impact <- press.impact(mmod,perturb=c(Fisheries=1))
  impact <- press.impact(mmod,perturb=c(Fisheries=qls[j]))
  
  ## Use 100 simulations
  n.sims <- 100
  results <- 0
  i <- 0
  while(i < n.sims) {
    
    ## Randomly choose edges to retain
    z <- s$select(runif(1))
    ## Sample community matrix
    W <- s$community()
    
    ## Check press condition and stability
    if(!(press(W) && stable.community(W))) next
    
    ## Monitor impact post press
    imp <- impact(W)
    imp[abs(imp)<1e-07]=0 #added to remove small perturbations
    results <- results + outer(sign(imp),-1:1,'==')
    i <- i+1
    
    rownames(results) <- levels(mmod$From)
    colnames(results) <- c('-','0','+')
    
    #write.csv(results, "./QPRESS_2020/SNS/SNS_fisheries_increase.csv")
  }
  
  
  #write simulations to .csv file
  write.csv(results, paste0("./output/QPRESS_2020/SNS_", scenarios[j],".csv"))
  
  
  #bar plot
    library(RColorBrewer)
    pal <- brewer.pal(n=5,"Greys")[5:1]
    
    prop <- results/rowSums(results)
    r <- colSums(t(prop)*(-1:1)) #this is the number we need to extract to the analysis
    #filename <-  ("./QPRESS_2020/SNS_fisheries_increase.pdf")
    filename <-  paste0("./output/QPRESS_2020/SNS/SNS_", scenarios[j],".pdf")
    pdf(filename)
    #adjust barplot if you want it ordered by proportion
    par(mar=c(5,10,1,1)+0.1)
   # barplot(t(prop),
    #        horiz=T,cex.names=0.7,cex.axis=0.8,las=2,border=T, col=pal,xlab="Proportion")
    
    opar <- par(mar=c(5,10,1,1)+0.1)
    barplot(t(prop[order(r),]),
          horiz=T,cex.names=0.6,cex.axis=0.8,las=2,border=T,col=pal,xlab="Proportion")
   dev.off()
    #par(opar)

}

```
##### SNS single output tables 
```{r SNS single output table}

tfiles <- list.files("./output/QPRESS_2020/SNS/", pattern=".csv")
i <- 1
results <- read_csv(paste0("./output/QPRESS_2020/SNS/", tfiles[i]))[,1]


for (i in 1:length(tfiles)){ 
tdf <- read_csv(paste0("./output/QPRESS_2020/SNS/", tfiles[i]))

  colnames(tdf) <- c("Variable", paste0(substr(tfiles[i], 5, 22), "_Down"), paste0(substr(tfiles[i], 5, 22), "_0"), paste0(substr(tfiles[i], 5, 22), "_Up") )
  results <- as.data.table(cbind(results, tdf[2:4]))
}

results$Fisheries_decrease <- results$fisheries_decrease_Up-results$fisheries_decrease_Down
results$Fisheries_increase <- results$fisheries_increase_Up-results$fisheries_increase_Down

write_csv(results, "./output/QPRESS_2020/SNS/SNS_QPRESS_All_scenarios.csv")
```



#### Skagerrak
only changed demersal fisheries to match the EwE scenarios. 
```{r Skagerrak QPRESS scenarios to CSV & barplot, echo=FALSE, warning=FALSE}

mmod <- skag_el

scenarios_f <- c("fisheries_incr", "fisheries_decr" )
scenarios_m <- c("marmam_incr", "marmam_decr")
qls <- c(1,-1)

A <- adjacency.matrix(mmod)

## Function to generate the community matrix
s <- community.sampler(mmod)

#Fisheries Scenarios
for (j in 1:length(scenarios_f)){ 
  ## Function to check the validation condition
  press <- press.validate(mmod, perturb=c(Fisheries_D=qls[j]), monitor=F)
  
  #press <- press.validate(mmod, perturb=c(Fisheries_D=qls[j], Fisheries_P=qls[j], Fisheries_Rec=qls[j]), monitor=F)
  #press <- press.validate(mmod, perturb=c(Fisheries_D=qls[j],  monitor=F)
  
  ## Function to define the perturbation scenario
  impact <- press.impact(mmod,perturb=c(Fisheries_D=qls[j]))
 # impact <- press.impact(mmod,perturb=c(Fisheries_D=qls[j], Fisheries_P=qls[j], Fisheries_Rec=qls[j]))
  
  #impact <- press.impact(mmod,perturb=c(Fisheries_D=qls[j]))
  
  ## Use 100 simulations
  n.sims <- 100
  results <- 0
  i <- 0
  while(i < n.sims) {
    
    ## Randomly choose edges to retain
    z <- s$select(runif(1))
    ## Sample community matrix
    W <- s$community()
    
    ## Check press condition and stability
    if(!(press(W) && stable.community(W))) next
    
    ## Monitor impact post press
    imp <- impact(W)
    imp[abs(imp)<1e-07]=0 #added to remove small perturbations
    results <- results + outer(sign(imp),-1:1,'==')
    i <- i+1
    
    rownames(results) <- levels(mmod$From)
    colnames(results) <- c('-','0','+')
    
  }
  
  #write simulations to .csv file
  write.csv(results, paste0("./output/QPRESS_2020/Skagerrak/Skagerrak_", scenarios_f[j],".csv"))
  
  #bar plot
    library(RColorBrewer)
    pal <- brewer.pal(n=5,"Greys")[5:1]
    
    prop <- results/rowSums(results)
    r <- colSums(t(prop)*(-1:1)) #this is the number we need to extract to the analysis
    filename <-  paste0("./output/QPRESS_2020/Skagerrak/Skagerrak_", scenarios_f[j],".pdf")
    pdf(filename)
    
    #adjust barplot if you want it ordered by proportion
    par(mar=c(5,10,1,1)+0.1)
    opar <- par(mar=c(5,10,1,1)+0.1)
    barplot(t(prop[order(r),]),
          horiz=T,cex.names=0.6,cex.axis=0.8,las=2,border=F,col=pal,xlab="Proportion")
   dev.off()
    #par(opar)

}


#marine mammal scenarios
for (j in 1:length(scenarios_m)){ 
  ## Function to check the validation condition
  press <- press.validate(mmod, perturb=c(Mammals=qls[j]), monitor=F)
  
  ## Function to define the perturbation scenario
  impact <- press.impact(mmod,perturb=c(Mammals=qls[j]))
  
  ## Use 100 simulations
  n.sims <- 100
  results <- 0
  i <- 0
  while(i < n.sims) {
    
    ## Randomly choose edges to retain
    z <- s$select(runif(1))
    ## Sample community matrix
    W <- s$community()
    
    ## Check press condition and stability
    if(!(press(W) && stable.community(W))) next
    
    ## Monitor impact post press
    imp <- impact(W)
    imp[abs(imp)<1e-07]=0 #added to remove small perturbations
    results <- results + outer(sign(imp),-1:1,'==')
    i <- i+1
    
    rownames(results) <- levels(mmod$From)
    colnames(results) <- c('-','0','+')
    
  }
  
  #write simulations to .csv file
  write.csv(results, paste0("./output/QPRESS_2020/Skagerrak/Skagerrak_", scenarios_m[j],".csv"))
  
  #bar plot
    library(RColorBrewer)
    pal <- brewer.pal(n=5,"Greys")[5:1]
    
    prop <- results/rowSums(results)
    r <- colSums(t(prop)*(-1:1)) #this is the number we need to extract to the analysis
    filename <-  paste0("./output/QPRESS_2020/Skagerrak/Skagerrak_", scenarios_m[j],".pdf")
    pdf(filename)
    
    #adjust barplot if you want it ordered by proportion
    par(mar=c(5,10,1,1)+0.1)
    opar <- par(mar=c(5,10,1,1)+0.1)
    barplot(t(prop[order(r),]),
          horiz=T,cex.names=0.6,cex.axis=0.8,las=2,border=F,col=pal,xlab="Proportion")
   dev.off()
    #par(opar)

}
```

##### Skagerrak single output tables
```{r Skagerrak single output table}

tfiles <- list.files("./output/QPRESS_2020/Skagerrak/", pattern=".csv")
i <- 1
results <- read_csv(paste0("./output/QPRESS_2020/Skagerrak/", tfiles[i]))[,1]


for (i in 1:length(tfiles)){ 
tdf <- read_csv(paste0("./output/QPRESS_2020/Skagerrak/", tfiles[i]))

  colnames(tdf) <- c("Variable", paste0(substr(tfiles[i], 11, 24), "_Down"), paste0(substr(tfiles[i], 11, 24), "_0"), paste0(substr(tfiles[i], 11, 24), "_Up") )
  results <- as.data.table(cbind(results, tdf[2:4]))
}

results$Fisheries_decrease <- results$fisheries_decr_Up-results$fisheries_decr_Down
results$Fisheries_increase <- results$fisheries_incr_Up-results$fisheries_incr_Down
results$marmam_decr <- results$marmam_decr.cs_Up-results$marmam_decr.cs_Down
results$marmam_incre <- results$marmam_incr.cs_Up-results$marmam_incr.cs_Down

write_csv(results, "./output/QPRESS_2020/Skagerrak/Skagerrak_QPRESS_All_scenarios.csv")
```


#### Kattegat
including only demersal fisheries
```{r Kattegat QPRESS scenarios to CSV & barplot, echo=FALSE, warning=FALSE}

mmod <- katt_el

scenarios_f <- c("fisheries_incr", "fisheries_decr" )
scenarios_m <- c("marmam_incr", "marmam_decr")
qls <- c(1,-1)

A <- adjacency.matrix(mmod)

## Function to generate the community matrix
s <- community.sampler(mmod)

#Fisheries Scenarios
for (j in 1:length(scenarios_f)){ 
  ## Function to check the validation condition
  press <- press.validate(mmod, perturb=c(Fisheries_Dem=qls[j]), monitor=F)
  #   press <- press.validate(mmod, perturb=c(Fisheries_Dem=qls[j], Fisheries_Pel=qls[j], Fisheries_Rec=qls[j]), monitor=F)
  
  ## Function to define the perturbation scenario
  impact <- press.impact(mmod,perturb=c(Fisheries_Dem=qls[j]))
  # impact <- press.impact(mmod,perturb=c(Fisheries_Dem=qls[j], Fisheries_Pel=qls[j], Fisheries_Rec=qls[j]))
  
  ## Use 100 simulations
  n.sims <- 100
  results <- 0
  i <- 0
  while(i < n.sims) {
    
    ## Randomly choose edges to retain
    z <- s$select(runif(1))
    ## Sample community matrix
    W <- s$community()
    
    ## Check press condition and stability
    if(!(press(W) && stable.community(W))) next
    
    ## Monitor impact post press
    imp <- impact(W)
    imp[abs(imp)<1e-07]=0 #added to remove small perturbations
    results <- results + outer(sign(imp),-1:1,'==')
    i <- i+1
    
    rownames(results) <- levels(mmod$From)
    colnames(results) <- c('-','0','+')
    
  }
  
  #write simulations to .csv file
  write.csv(results, paste0("./output/QPRESS_2020/Kattegat/Kattegat_", scenarios_f[j],".csv"))
  
  #bar plot
    library(RColorBrewer)
    pal <- brewer.pal(n=5,"Greys")[5:1]
    
    prop <- results/rowSums(results)
    r <- colSums(t(prop)*(-1:1)) #this is the number we need to extract to the analysis
    filename <-  paste0("./output/QPRESS_2020/Kattegat/Kattegat_", scenarios_f[j],".pdf")
    pdf(filename)
    
    #adjust barplot if you want it ordered by proportion
    par(mar=c(5,10,1,1)+0.1)
    opar <- par(mar=c(5,10,1,1)+0.1)
    barplot(t(prop[order(r),]),
          horiz=T,cex.names=0.5,cex.axis=0.8,las=2,border=F,col=pal,xlab="Proportion")
   dev.off()
    #par(opar)

}


#marine mammal scenarios
for (j in 1:length(scenarios_m)){ 
  ## Function to check the validation condition
  press <- press.validate(mmod, perturb=c(Seals=qls[j]), monitor=F)
  
  ## Function to define the perturbation scenario
  impact <- press.impact(mmod,perturb=c(Seals=qls[j]))
  
  ## Use 100 simulations
  n.sims <- 100
  results <- 0
  i <- 0
  while(i < n.sims) {
    
    ## Randomly choose edges to retain
    z <- s$select(runif(1))
    ## Sample community matrix
    W <- s$community()
    
    ## Check press condition and stability
    if(!(press(W) && stable.community(W))) next
    
    ## Monitor impact post press
    imp <- impact(W)
    imp[abs(imp)<1e-07]=0 #added to remove small perturbations
    results <- results + outer(sign(imp),-1:1,'==')
    i <- i+1
    
    rownames(results) <- levels(mmod$From)
    colnames(results) <- c('-','0','+')
    
  }
  
  #write simulations to .csv file
  write.csv(results, paste0("./output/QPRESS_2020/Kattegat/Kattegat_", scenarios_m[j],".csv"))
  
  #bar plot
    library(RColorBrewer)
    pal <- brewer.pal(n=5,"Greys")[5:1]
    
    prop <- results/rowSums(results)
    r <- colSums(t(prop)*(-1:1)) #this is the number we need to extract to the analysis
    filename <-  paste0("./output/QPRESS_2020/Kattegat/Kattegat_", scenarios_m[j],".pdf")
    pdf(filename)
    
    #adjust barplot if you want it ordered by proportion
    par(mar=c(5,10,1,1)+0.1)
    opar <- par(mar=c(5,10,1,1)+0.1)
    barplot(t(prop[order(r),]),
          horiz=T,cex.names=0.5,cex.axis=0.8,las=2,border=F,col=pal,xlab="Proportion")
   dev.off()
    #par(opar)

}
```


##### Kattegat single output tables
```{r Kattegat single output table}

tfiles <- list.files("./output/QPRESS_2020/Kattegat/", pattern=".csv")
i <- 1
results <- read_csv(paste0("./output/QPRESS_2020/Kattegat/", tfiles[i]))[,1]


for (i in 1:length(tfiles)){ 
tdf <- read_csv(paste0("./output/QPRESS_2020/Kattegat/", tfiles[i]))

  colnames(tdf) <- c("Variable", paste0(substr(tfiles[i], 10, 23), "_Down"), paste0(substr(tfiles[i], 10, 23), "_0"), paste0(substr(tfiles[i], 10, 23), "_Up") )
  results <- as.data.table(cbind(results, tdf[2:4]))
}

results$Fisheries_decrease <- results$fisheries_decr_Up-results$fisheries_decr_Down
results$Fisheries_increase <- results$fisheries_incr_Up-results$fisheries_incr_Down
results$marmam_decr <- results$marmam_decr.cs_Up-results$marmam_decr.cs_Down
results$marmam_incre <- results$marmam_incr.cs_Up-results$marmam_incr.cs_Down

write_csv(results, "./output/QPRESS_2020/Kattegat/Kattegat_QPRESS_All_scenarios.csv")
```

#### Norwegian Trench
Demersal fisheries only
```{r  NorwTrench QPRESS scenarios to CSV & barplot, echo=FALSE, warning=FALSE}

mmod <- Nor_el

scenarios_f <- c("fisheries_incr", "fisheries_decr" )

qls <- c(1,-1)

A <- adjacency.matrix(mmod)

## Function to generate the community matrix
s <- community.sampler(mmod)

#Fisheries Scenarios
for (j in 1:length(scenarios_f)){ 
  ## Function to check the validation condition
  press <- press.validate(mmod, perturb=c(Fisheries_ind=qls[j], L_mesh_trawling=qls[j], Fisheries_shrimp=qls[j]), monitor=F) 
  # press <- press.validate(mmod, perturb=c(Fisheries_ind=qls[j], L_mesh_trawling=qls[j], Fisheries_pel=qls[j],  Fisheries_shrimp=qls[j]), monitor=F)
  
  ## Function to define the perturbation scenario
  impact <- press.impact(mmod,perturb=c(Fisheries_ind=qls[j], L_mesh_trawling=qls[j], Fisheries_shrimp=qls[j]))
  # impact <- press.impact(mmod,perturb=c(Fisheries_ind=qls[j], L_mesh_trawling=qls[j], Fisheries_pel=qls[j],  Fisheries_shrimp=qls[j]))
  
  ## Use 100 simulations
  n.sims <- 100
  results <- 0
  i <- 0
  while(i < n.sims) {
    
    ## Randomly choose edges to retain
    z <- s$select(runif(1))
    ## Sample community matrix
    W <- s$community()
    
    ## Check press condition and stability
    if(!(press(W) && stable.community(W))) next
    
    ## Monitor impact post press
    imp <- impact(W)
    imp[abs(imp)<1e-07]=0 #added to remove small perturbations
    results <- results + outer(sign(imp),-1:1,'==')
    i <- i+1
    
    rownames(results) <- levels(mmod$From)
    colnames(results) <- c('-','0','+')
    
  }
  
  #write simulations to .csv file
  write.csv(results, paste0("./output/QPRESS_2020/NorwTrench/NorwTrench_", scenarios_f[j],".csv"))
  
  #bar plot
    library(RColorBrewer)
    pal <- brewer.pal(n=5,"Greys")[5:1]
    
    prop <- results/rowSums(results)
    r <- colSums(t(prop)*(-1:1)) #this is the number we need to extract to the analysis
    filename <-  paste0("./output/QPRESS_2020/NorwTrench/NorwTrench_", scenarios_f[j],".pdf")
    pdf(filename)
    
    #adjust barplot if you want it ordered by proportion
    par(mar=c(5,10,1,1)+0.1)
    opar <- par(mar=c(5,10,1,1)+0.1)
    barplot(t(prop[order(r),]),
          horiz=T,cex.names=0.6,cex.axis=0.8,las=2,border=F,col=pal,xlab="Proportion")
   dev.off()
    #par(opar)

}

```

##### Norw Trench single output table
```{r NorwTrench single output table}

tfiles <- list.files("./output/QPRESS_2020/NorwTrench/", pattern=".csv")
i <- 1
results <- read_csv(paste0("./output/QPRESS_2020/NorwTrench/", tfiles[i]))[,1]


for (i in 1:length(tfiles)){ 
tdf <- read_csv(paste0("./output/QPRESS_2020/NorwTrench/", tfiles[i]))

  colnames(tdf) <- c("Variable", paste0(substr(tfiles[i], 12, 25), "_Down"), paste0(substr(tfiles[i], 12, 25), "_0"), paste0(substr(tfiles[i], 12, 25), "_Up") )
  results <- as.data.table(cbind(results, tdf[2:4]))
}

results$Fisheries_decrease <- results$fisheries_decr_Up-results$fisheries_decr_Down
results$Fisheries_increase <- results$fisheries_incr_Up-results$fisheries_incr_Down

write_csv(results, "./output/QPRESS_2020/NorwTrench/NorwTrench_QPRESS_All_scenarios.csv")
```



