---
title: "QPRESS EwE comparison"
author: "Erik Olsen"
date: "10/13/2020"
output: html_document
---
## Description
This script compares the output from the scenarios run on the Kattegat and NS EwE models with the scenarios run using QPRESS for the mental models of the Southern North Sea, Kattegat, Skagerrak and Norwegian Trench.   


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Libraries
```{r libraries echo=FALSE, warning=FALSE}
library("tidyverse")
library("readxl")
library(grid) 
library(gridExtra)
library(patchwork)
library(devtools) 
#install_github("vqv/ggbiplot", force = TRUE)
library(ggbiplot)

library(flextable)
library(scales)
library(webshot)
webshot::install_phantomjs()
library(ggstatsplot)
```

## Load data files
```{r  load files echo=FALSE, warning=FALSE}
Katt_EwE_bio <- read_csv("./figs/biomass Kattegat.csv")
Katt_EwE_catch <- read_csv("./catches Kattegat.csv")
NS_EwE_bio <- read_csv("./figs/biomass NS.csv")
NS_EwE_catch <- read_csv("./figs/catch NS.csv")
SNS_QP <- read_csv("./output/QPRESS_2020/SNS/SNS_QPRESS_All_scenarios.csv")
Katt_QP <- read_csv("./output/QPRESS_2020/Kattegat/Kattegat_QPRESS_All_scenarios.csv")
Skag_QP <- read_csv("./output/QPRESS_2020/Skagerrak/Skagerrak_QPRESS_All_scenarios.csv")
Nor_QP <- read_csv("./output/QPRESS_2020/NorwTrench/NorwTrench_QPRESS_All_scenarios.csv")
```
## Create joint data sets for the Southern North Sea, Skagerrak, Norwegian Trench and Kattegat
First need to make comparable groups of the model variables. 

### Southern North Sea combined data set
Combine SNS_QP, NS_EwE_bio and NS_EwE_catch  
  
SNS_QP model only has "benthic_hab", "biol_spec" and "fisheries" that match variables in the NS_EwE model, so all catches in the NS_EwE model are mapped to "fisheries" and the biological components are split and mapped to "benthic_hab" and "biol_spec" components.  
  
  
"benthic_hab" consisist of the EwE components: 
* "Small.infauna..polychaetes."  
* "Sessile.epifauna" 
* "Meiofauna"   
* "Benthic.microflora..incl.Bacteria.protozoa.."
* "Detritus...POM...sediment"   
  
All other EwE biological component are mapped to "biol_spec"  
  
When combining groups a simple average in the normalized responses are calculated, which may be incorrect as small and large groups are given equal weight. 


```{r SNS combined, echo=FALSE, warning=FALSE}
# get list of variable names
NS_EwE_bio$scenario
NS_EwE_catch$scenario
SNS_QP$X1

#######
#benthos / benthic habitat group
EWEbenthos <-NS_EwE_bio %>% filter(scenario%in%c("Small.infauna..polychaetes.",  "Sessile.epifauna", "Meiofauna", "Benthic.microflora..incl.Bacteria.protozoa..", "Detritus...POM...sediment")) %>% summarise_each(funs(mean))

QP_benthos <- SNS_QP %>% filter(X1=="Benthic_Hab") 

tqp <- as.data.frame(t(QP_benthos[,8:9]))
tqp[,2] <- c("MM")
tqp$V1<-tqp$V1/100+1
tewe <- as.data.frame(t(EWEbenthos[,3:6]))
tewe[,2]<-c("EwE")

SNSbenthos <- rbind(tqp, tewe)
SNSbenthos$scenario <- rownames(SNSbenthos)
colnames(SNSbenthos) <- c("Response", "model", "Scenario")

SNSbenthos$Scenario <- factor(SNSbenthos$Scenario,levels = c("No fishing","-50%  fishing","+25%  fishing","+75%  fishing", "Fisheries_decrease", "Fisheries_increase"))


#######
#other biological groups
EWEbio <-   NS_EwE_bio %>% filter(!scenario%in%c("Small.infauna..polychaetes.",  "Sessile.epifauna", "Meiofauna", "Benthic.microflora..incl.Bacteria.protozoa..", "Detritus...POM...sediment")) %>% summarise_each(funs(mean))

QPbio <- SNS_QP %>% filter(X1=="Biol_Spec") 

tqp2 <- as.data.frame(t(QPbio[,8:9]))
tqp2[,2] <- c("MM")
tqp2$V1<-tqp2$V1/100+1
tewe2 <- as.data.frame(t(EWEbio[,3:6]))
tewe2[,2]<-c("EwE")

SNSbio <- rbind(tqp2, tewe2)
SNSbio$scenario <- rownames(SNSbio)
colnames(SNSbio) <- c("Response", "model", "Scenario")

SNSbio$Scenario <- factor(SNSbio$Scenario,levels = c("No fishing","-50%  fishing","+25%  fishing","+75%  fishing", "Fisheries_decrease", "Fisheries_increase"))


#####
# catch data

EWEcatch<-NS_EwE_catch %>% summarise_each(mean)

QPfish <- SNS_QP %>% filter(X1=="Fisheries") 

tqp3 <- as.data.frame(t(QPfish[,8:9]))
tqp3[,2] <- c("MM")
tqp3$V1<-tqp3$V1/100+1
tewe3 <- as.data.frame(t(EWEcatch[,3:6]))
tewe3[,2]<-c("EwE")

SNScatch <- rbind(tqp3, tewe2)
SNScatch$scenario <- rownames(SNScatch)
colnames(SNScatch) <- c("Response", "model", "Scenario")

SNScatch$Scenario <- factor(SNScatch$Scenario,levels = c("No fishing","-50%  fishing","+25%  fishing","+75%  fishing", "Fisheries_decrease", "Fisheries_increase"))


```

####Southern North Sea comparison plots
```{r SNS comp plots, echo=FALSE, warning=FALSE}

SNSbenthos$component <- c("benthos")
SNSbio$component <- c("biological components")
SNScatch$component <- c("catch")

SNS_all <- rbind(SNSbenthos, SNSbio, SNScatch)

SNS_all$Scenario <- factor(SNS_all$Scenario,levels = c("No fishing","-50%  fishing", "Fisheries_decrease", "+25%  fishing","+75%  fishing",  "Fisheries_increase"))

#plot all 
ggplot(SNS_all, aes(Scenario,Response-1, fill=model)) +geom_col(aes(y=Scenario, x=Response-1)) +theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + theme(axis.title.y = element_blank(), axis.title.x = element_blank()) + ggtitle("Southern North Sea") +facet_wrap(~component, scales="free_y", dir="v")

ggsave("./figs/SNS_all.png", width=14, height=10, units="cm")

# benthos plot
#ggplot(SNSbenthos, aes(Scenario,Response-1, fill=model)) +geom_col() +theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + theme(axis.title.y = element_blank()) + ggtitle("Southern North Sea: Benthos/habitats")


```

#### SNS combined table

```{r comp table Skagerrak, echo=FALSE, warning=FALSE}
Nor_comb2 <- SNS_all
Nor_comb2$Response <- Nor_comb2$Response-1
Nor_wide <- Nor_comb2 %>% pivot_wider(names_from = c(Scenario, model), values_from=Response) #exclude mackerel (not in MM data)
Nor_wide <- Nor_wide[,c(1,2,4,5, 3,6,7)]
#colnames(Nor_wide) <- c("component", "Fisheries_decrease_MM", "No_fishing_EwE", "E-50% _fishing_EwE","Fisheries_increase_MM","E+25%_fishing_EwE", "E+75%_fishing_EwE")    

scale <- scales::col_numeric(domain= c(-1.75, 1.75), palette ="PiYG")  
ft_2 <- flextable(Nor_wide)
  ft_2 <- bg(ft_2, j = colnames(Nor_wide)[2:7], bg = scale)
  ft_2 <- set_formatter_type(ft_2)
  ft_2  
  
ft_2 <- add_header_row(ft_2,  colwidths = c(1, 3, 3),  values = c(" ", "Decrease fisheries", "Increase fisheries"))

ft_2 <- set_header_labels(ft_2, component = " ", 
    Fisheries_decrease_MM = "MM: decrease", "No fishing_EwE" = "EwE: -100%",
    "-50%  fishing_EwE" = "EwE: -50%", Fisheries_increase_MM = "MM: increase", "+25%  fishing_EwE" = "EwE: +25%", "+75%  fishing_EwE" = "EwE: +75%")

ft_2


#export as docx and png
save_as_docx("SNS Mental model and EwE comparisons" = ft_2, path = "./figs/SNS_table.docx")
save_as_image(ft_2, path = "./figs/SNS_table.png")
```
#### SNS Correlation plot
```{r SNS Correlation, echo=FALSE}
set.seed(123)

ggcorrmat(
  data = Nor_wide, # data from which variable is to be taken
  cor.vars = 2:4, # specifying correlation matrix variables
  #type="nonparametric"
)

ggsave("./QPRESS_2020/figs/SNS_F_decr_corr_plot.png")


ggcorrmat(
  data = Nor_wide, # data from which variable is to be taken
  cor.vars = 5:7, # specifying correlation matrix variables
  #type="nonparametric"
)

ggsave("./figs/SNS_F_incr_corr_plot.png")

```



### Kattegat combined data sets
```{r Kattegat combined data, echo=FALSE, warning=FALSE}

# get list of variable names
#Katt_EwE_bio$scenario
#Katt_EwE_catch$group_n
#Katt_QP$X1

colnames(Katt_QP)[1] <- c("component")
colnames(Katt_EwE_bio)[2] <- c("component")
colnames(Katt_EwE_catch)[2] <- c("component")

# EWE Biol data
#map EwE names to MM names
kmapping <- read_excel("./Kattegat MM EwE mapping.xlsx")
Katt_EwE_bio$component_m <- kmapping$Mmmapping[match(Katt_EwE_bio$component, kmapping$EwE)]

Katt_QP$component_e <- c("Not_in_MM")
Katt_QP$component_e <- kmapping$EwE[match(Katt_QP$component, kmapping$Mmmapping)]
  

#convert til long format
Kbiol_long <- Katt_EwE_bio[,c(3:9)] %>% pivot_longer(!component_m, names_to = "scenario", values_to = "value", names_repair="unique")

Kbiol_long$model <- c("EwE")
colnames(Kbiol_long)[1] <- c("component")

#sum same biol components
Kbiol_long <- Kbiol_long %>% group_by(component, scenario) %>% dplyr::summarise(value=mean(value))
Kbiol_long$model <- c("EwE")

#EWE Catch data
Kec <- Katt_EwE_catch[,c(2:8)] %>% summarise_all(funs(mean))
Kec[,1] <- as.character(Kec[,1])
Kec[1,1] <- c("Fisheries_catch")
#to long format
Kec <- Kec %>% pivot_longer(!component, names_to = "scenario", values_to = "value", names_repair="unique") 
Kec$model <- c("EwE")



## MM data
#biol components
Kqp_long <-  Katt_QP[!is.na(Katt_QP$component_e),][,c(1,14:17)] %>% pivot_longer(!component, names_to = "scenario", values_to = "value", names_repair="unique")
Kqp_long$model <- c("MM")
#catch all fisheries
Kfish <- subset(Katt_QP,component == "Fisheries_Dem" | component == "Fisheries_Pel" |  component == "Fisheries_Rec")[,c(1,14:17)] %>% pivot_longer(!component, names_to = "scenario", values_to = "value", names_repair="unique") %>% group_by(scenario) %>% dplyr::summarise(value=mean(value))
Kfish$component <- c("Fisheries_catch")
Kfish <- Kfish[,c(3,1,2)]
Kfish$model <- c("MM")


#combine biol and catch
Kqp_long <- rbind(Kqp_long, Kfish)

#normalize all data
Kqp_long$value <- Kqp_long$value/100+1
Kqp_long$model <- c("MM")

#Combine data
Katt_comb <- rbind(Kqp_long, Kbiol_long, Kec)


Katt_comb$scenario <- factor(Katt_comb$scenario,levels = c("No fishing","-50% demersal fishing","Fisheries_decrease", "+25% demersal fishing","+75% demersal fishing",   "Fisheries_increase", "seal biomass X 10","seal biomass X 2","marmam_decr", "marmam_incre"))



```


#### Kattegat comparison plots
```{r Kattegat comparison plots, echo=FALSE, warning=FALSE}
# all components
ggplot(Katt_comb, aes(scenario,value-1, fill=model)) +geom_col(aes(y=scenario, x=value-1)) +theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + theme(axis.title.y = element_blank(), axis.title.x = element_blank()) + ggtitle("Kattegat") +facet_wrap(vars(component), scales="free_x")



ggsave("./figs/Kattegat_bio_catch.png", height=18, width=14, units="cm")

```

#### Kattegat combined table

```{r comp table Kattegat, echo=FALSE, warning=FALSE}
Nor_comb2 <- Katt_comb
Nor_comb2$value <- Nor_comb2$value-1
Nor_wide <- Nor_comb2 %>% pivot_wider(names_from = c(scenario, model), values_from=value) #exclude mackerel (not in MM data)
Nor_wide <- Nor_wide[,c(1,2,9,6, 3,7,8, 4, 5, 10,11)]
#colnames(Nor_wide) <- c("component", "Fisheries_decrease_MM", "No_fishing_EwE", "E-50% _fishing_EwE","Fisheries_increase_MM","E+25%_fishing_EwE", "E+75%_fishing_EwE")    

scale <- scales::col_numeric(domain= c(-0.8, 0.8), palette ="PiYG")  
ft_2 <- flextable(Nor_wide)
  ft_2 <- bg(ft_2, j = colnames(Nor_wide)[2:7], bg = scale)
  ft_2 <- set_formatter_type(ft_2)
  ft_2  
  
  scale2 <- scales::col_numeric(domain= c(-9.01, 9.01), palette ="PiYG")  
  ft_2 <- bg(ft_2, j = colnames(Nor_wide)[8:11], bg = scale2)
  ft_2
  
ft_2 <- add_header_row(ft_2,  colwidths = c(1, 3, 3, 4),  values = c(" ", "Decrease fisheries", "Increase fisheries", "Seals"))

ft_2 <- set_header_labels(ft_2, component = " ", 
    Fisheries_decrease_MM = "MM: decrease", "No fishing_EwE" = "EwE: -100%",
    "-50% demersal fishing_EwE" = "EwE: -50%", Fisheries_increase_MM = "MM: increase", "+25% demersal fishing_EwE" = "EwE: +25%", "+75% demersal fishing_EwE" = "EwE: +75%", "marmam_decr_MM" ="MM: decrease", "marmam_incre_MM" = "MM: increase", "seal biomass X 10_EwE" = "EwE: seals  X 10", "seal biomass X 2_EwE" = "EwE: Seals X 2"  )

ft_2


#export as docx and png
save_as_docx("Kattegat Mental model and EwE comparisons" = ft_2, path = "./figs/Kattegat_table.docx")
save_as_image(ft_2, path = "./figs/Kattegat_table.png")
```

#### Kattegat correlation plot
```{r kattegat Correlation, echo=FALSE}

ggcorrmat(
  data = Nor_wide, # data from which variable is to be taken
  cor.vars = 2:4, # specifying correlation matrix variables
  #type="nonparametric"
)

ggsave("./figs/kattegat_F_decr_corr_plot.png")

ggcorrmat(
  data = Nor_wide, # data from which variable is to be taken
  cor.vars = 5:7, # specifying correlation matrix variables
 # type="nonparametric"
)

ggsave("./figs/kattegat_F_incr_corr_plot.png")


ggcorrmat(
  data = Nor_wide, # data from which variable is to be taken
  cor.vars = c(9,10,11), # specifying correlation matrix variables
 # type="nonparametric"
)

ggsave("./figs/kattegat_seals_corr_plot.png")

```



### Skagerrak combined data sets
```{r Skagerrak combined data, echo=FALSE, warning=FALSE}
# get list of variable names


NS_EwE_bio$scenario
NS_EwE_catch$scenario
Skag_QP$X1

colnames(Skag_QP)[1] <- c("component")
colnames(NS_EwE_bio)[2] <- c("component")
colnames(NS_EwE_catch)[2] <- c("component")

# EWE Biol data
#map EwE names to MM names
kmapping <- read_excel("./data/NS_EwE/NS_skagerrak_mapping.xlsx")
NS_EwE_bio$component_m <- kmapping$Mmmapping[match(NS_EwE_bio$component, kmapping$EwE)]

Skag_QP$component_e <- c("Not_in_MM")
Skag_QP$component_e <- kmapping$EwE[match(Skag_QP$component, kmapping$Mmmapping)]
  

#convert til long format
Sbiol_long <- NS_EwE_bio[,c(3:7)] %>% pivot_longer(!component_m, names_to = "scenario", values_to = "value", names_repair="unique")

Sbiol_long$model <- c("EwE")
colnames(Sbiol_long)[1] <- c("component")

#sum same biol components
Sbiol_long <- subset(Sbiol_long, component!="NoMMgroup")
Sbiol_long <- Sbiol_long %>% dplyr::group_by(component, scenario) %>% dplyr::summarise(value=mean(value))
Sbiol_long$model <- c("EwE")



#EWE Catch data
NS_EwE_catch$component_m <- kmapping$Mmmapping[match(NS_EwE_catch$component, kmapping$EwE)]

Kec <- subset(NS_EwE_catch,component_m!="NoMMgroup" )[,c(2:6)] %>% summarise_all(funs(mean))
Kec[,1] <- as.character(Kec[,1])
Kec[1,1] <- c("Fisheries_catch")
#to long format
Kec <- Kec %>% pivot_longer(!component, names_to = "scenario", values_to = "value", names_repair="unique") 
Kec$model <- c("EwE")



## MM data
#biol components
Kqp_long <-  Skag_QP[!is.na(Skag_QP$component_e),][,c(1,14:17)] %>% pivot_longer(!component, names_to = "scenario", values_to = "value", names_repair="unique")
Kqp_long$model <- c("MM")
#catch all fisheries
Kfish <- subset(Skag_QP,component == "Fisheries_D" | component == "Fisheries_P" |  component == "Fisheries_Rec")[,c(1,14:17)] %>% pivot_longer(!component, names_to = "scenario", values_to = "value", names_repair="unique") %>% dplyr::group_by(scenario) %>% dplyr::summarise(value=mean(value))
Kfish$component <- c("Fisheries_catch")
Kfish <- Kfish[,c(3,1,2)]
Kfish$model <- c("MM")


#combine biol and catch
Kqp_long <- rbind(Kqp_long, Kfish)

#normalize all data
Kqp_long$value <- Kqp_long$value/100+1
Kqp_long$model <- c("MM")

#Combine data
Skag_comb <- rbind(Kqp_long, Sbiol_long, Kec)


Skag_comb$scenario <- factor(Skag_comb$scenario,levels = c("No fishing","-50%  fishing","Fisheries_decrease", "+25%  fishing","+75%  fishing",   "Fisheries_increase", "marmam_decr", "marmam_incre"))


```

#### Skagerrak comparison plots
```{r Skagerrak comparison plots, echo=FALSE, warning=FALSE}

sc2 <- subset(Skag_comb, (scenario!="marmam_decr" & scenario!="marmam_incre") )
  
ggplot(sc2, aes(scenario,value-1, fill=model)) +geom_col(aes(y=scenario, x=value-1)) +theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + theme(axis.title.y = element_blank(), axis.title.x = element_blank()) + ggtitle("Skagerrak") +facet_wrap(vars(component), scales="free_x")

ggsave("./figs/Skagerrak_bio_catch.png", height=18, width=14, units="cm")

```
#### Skagerrak combined table

```{r comp table Skagerrak, echo=FALSE, warning=FALSE}
Nor_comb2 <- Skag_comb
Nor_comb2$value <- Nor_comb2$value-1
Nor_wide <- Nor_comb2 %>% pivot_wider(names_from = c(scenario, model), values_from=value) #exclude mackerel (not in MM data)
#Nor_wide <- Nor_wide[,c(1,2,9,6,3,7,8, 4, 5)]
# Without marine mammal scenarios
Nor_wide <- Nor_wide[,c(1,2,9,6,3,7,8)]
#colnames(Nor_wide) <- c("component", "Fisheries_decrease_MM", "No_fishing_EwE", "E-50% _fishing_EwE","Fisheries_increase_MM","E+25%_fishing_EwE", "E+75%_fishing_EwE")    

scale <- scales::col_numeric(domain= c(-1.14, 1.14), palette ="PiYG")  
ft_2 <- flextable(Nor_wide)
  ft_2 <- bg(ft_2, j = colnames(Nor_wide)[2:7], bg = scale)
  ft_2 <- set_formatter_type(ft_2)
  ft_2  
  
ft_2 <- add_header_row(ft_2,  colwidths = c(1, 3, 3),  values = c(" ", "Decrease fisheries", "Increase fisheries"))

#ft_2 <- set_header_labels(ft_2, component = " ", 
#    Fisheries_decrease_MM = "MM: decrease", "No fishing_EwE" = "EwE: -100%",
#    "-50%  fishing_EwE" = "EwE: -50%", Fisheries_increase_MM = "MM: increase", "+25%  fishing_EwE" = "EwE: +25%", "+75%  fishing_EwE" = "EwE: +75%", "marmam_decr_MM" ="MM: decrease", "marmam_incre_MM" = "MM: increase"  )

ft_2 <- set_header_labels(ft_2, component = " ", 
    Fisheries_decrease_MM = "MM: decrease", "No fishing_EwE" = "EwE: -100%",
    "-50%  fishing_EwE" = "EwE: -50%", Fisheries_increase_MM = "MM: increase", "+25%  fishing_EwE" = "EwE: +25%", "+75%  fishing_EwE" = "EwE: +75%" )

ft_2


#export as docx and png
save_as_docx("Skagerrak. Mental model and EwE comparisons" = ft_2, path = "./figs/Skagerrak_table.docx")
save_as_image(ft_2, path = "./figs/Skagerrak_table.png")
```
#### Skagerrak correlation plots
```{r skagerrak correlation plots}

ggcorrmat(
  data = Nor_wide, # data from which variable is to be taken
  cor.vars = 2:4, # specifying correlation matrix variables
  #type="nonparametric"
)

ggsave("./figs/skagerrak_F_decr_corr_plot.png")

ggcorrmat(
  data = Nor_wide, # data from which variable is to be taken
  cor.vars = 5:7, # specifying correlation matrix variables
 # type="nonparametric"
)

ggsave("./figs/skagerrak_F_incr_corr_plot.png")


```



### Norwegian trench combined data sets
```{r Nowegian Trench combined data, echo=FALSE, warning=FALSE}
#NS_EwE_bio$scenario
#NS_EwE_catch$scenario
Nor_QP$X1

colnames(Nor_QP)[1] <- c("component")
colnames(NS_EwE_bio)[2] <- c("component")
colnames(NS_EwE_catch)[2] <- c("component")

# EWE Biol data
#map EwE names to MM names
kmapping <- read_excel("./NS_NorTrench_mapping.xlsx")
NS_EwE_bio$component_m <- kmapping$Mmmapping[match(NS_EwE_bio$component, kmapping$EwE)]

Nor_QP$component_e <- c("Not_in_MM")
Nor_QP$component_e <- kmapping$EwE[match(Nor_QP$component, kmapping$Mmmapping)]
  

#convert til long format
Sbiol_long <- NS_EwE_bio[,c(3:7)] %>% pivot_longer(!component_m, names_to = "scenario", values_to = "value", names_repair="unique")

Sbiol_long$model <- c("EwE")
colnames(Sbiol_long)[1] <- c("component")

#sum same biol components
Sbiol_long <- subset(Sbiol_long, component!="NoMMgroup")
Sbiol_long <- Sbiol_long %>% group_by(component, scenario) %>% dplyr::summarise(value=mean(value))
Sbiol_long$model <- c("EwE")



#EWE Catch data
NS_EwE_catch$component_m <- kmapping$Mmmapping[match(NS_EwE_catch$component, kmapping$EwE)]

Kec <- subset(NS_EwE_catch,component_m!="NoMMgroup" )[,c(2:6)] %>% dplyr::summarise_all(funs(mean))
Kec[,1] <- as.character(Kec[,1])
Kec[1,1] <- c("Fisheries_catch")
#to long format
Kec <- Kec %>% pivot_longer(!component, names_to = "scenario", values_to = "value", names_repair="unique") 
Kec$model <- c("EwE")



## MM data
#biol components
Kqp_long <-  Nor_QP[!is.na(Nor_QP$component_e),][,c(1,8:9)] %>% pivot_longer(!component, names_to = "scenario", values_to = "value", names_repair="unique")
Kqp_long$model <- c("MM")
#catch all fisheries
Kfish <- subset(Nor_QP,component == "Fisheries_ind" | component == "Fisheries_pel" |  component == "Fisheries_shrimp" |  component == "L_mesh_trawling")[,c(1,8:9)] %>% pivot_longer(!component, names_to = "scenario", values_to = "value", names_repair="unique") %>% group_by(scenario) %>% dplyr::summarise(value=mean(value))
Kfish$component <- c("Fisheries_catch")
Kfish <- Kfish[,c(3,1,2)]
Kfish$model <- c("MM")


#combine biol and catch
Kqp_long <- rbind(Kqp_long, Kfish)

#normalize all data
Kqp_long$value <- Kqp_long$value/100+1
Kqp_long$model <- c("MM")

#Combine data
Nor_comb <- rbind(Kqp_long, Sbiol_long, Kec)


Nor_comb$scenario <- factor(Nor_comb$scenario,levels = c("No fishing","-50%  fishing","Fisheries_decrease", "+25%  fishing","+75%  fishing",   "Fisheries_increase"))


```

#### Norwegian trench comparison plots
```{r Norwegian trench comparison plots, echo=FALSE, warning=FALSE}
ggplot(subset(Nor_comb, component != "Mackerel"), aes(scenario,value-1, fill=model)) +geom_col(aes(y=scenario, x=value-1)) +theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + theme(axis.title.y = element_blank(), axis.title.x = element_blank()) + ggtitle("Norwegian Trench") +facet_wrap(~component, scales="free_x")

ggsave("./figs/NorwTrench_bio_catch.png", height=12, width=14, unit="cm")
```


### Table of model responses

```{r comp table Norw trenc, echo=FALSE, warning=FALSE}
Nor_comb2 <- Nor_comb
Nor_comb2$value <- Nor_comb2$value-1
Nor_wide <- subset(Nor_comb2, component != "Mackerel") %>% pivot_wider(names_from = c(scenario, model), values_from=value) #exclude mackerel (not in MM data)
Nor_wide <- Nor_wide[,c(1,2,7,4,3,5,6)]
#colnames(Nor_wide) <- c("component", "Fisheries_decrease_MM", "No_fishing_EwE", "E-50% _fishing_EwE","Fisheries_increase_MM","E+25%_fishing_EwE", "E+75%_fishing_EwE")    

scale <- scales::col_numeric(domain= c(-2.21, 2.21), palette ="PiYG")  
ft_2 <- flextable(Nor_wide)
  ft_2 <- bg(ft_2, j = colnames(Nor_wide)[2:7], bg = scale)
  ft_2 <- set_formatter_type(ft_2)
  ft_2  
  
ft_2 <- add_header_row(ft_2,  colwidths = c(1, 3, 3),  values = c(" ", "Decrease fisheries", "Increase fisheries"))

ft_2 <- set_header_labels(ft_2, component = " ", 
    Fisheries_decrease_MM = "MM: decrease", "No fishing_EwE" = "EwE: -100%",
    "-50%  fishing_EwE" = "EwE: -50%", Fisheries_increase_MM = "MM: increase", "+25%  fishing_EwE" = "EwE: +25%", "+75%  fishing_EwE" = "EwE: +75%" )

ft_2


#export as docx and png
save_as_docx("Norwegian Trench Mental model and EwE comparisons" = ft_2, path = "./figs/NorwTrench_table.docx")
save_as_image(ft_2, path = "./figs/NorwTrench_table.png")
```

#### Norwegian Trench radar plots
```{r norw trench radar, echo=FALSE, warning=FALSE}
nw_t <- t(Nor_wide[,2:7])
colnames(nw_t) <- Nor_wide$component

nw_t <- rbind(rep(1,10) , rep(-1,10) , nw_t)


png("radar_norwtrench.png")
radarchart(as.data.frame(nw_t))

legend(
  x = "topright", legend = rownames(nw_t[-c(1,2),]), horiz = TRUE,
  bty = "n", pch = 20 ,
  text.col = "black", cex = 1, pt.cex = 1.5
  )
dev.off()

getwd()

```


#### Norwegian trench correlation plots
```{r norw trench correlation plots}

ggcorrmat(
  data = Nor_wide, # data from which variable is to be taken
  cor.vars = 2:4, # specifying correlation matrix variables
  #type="nonparametric"
)

ggsave("./figs/NorwTrench_F_decr_corr_plot.png")

ggcorrmat(
  data = Nor_wide, # data from which variable is to be taken
  cor.vars = 5:7, # specifying correlation matrix variables
 # type="nonparametric"
)

ggsave("./figs/NorwTrench_F_incr_corr_plot.png")

```


## PCA analysis of EwE vs MM results
Cannot carry out a PCA on all data combined because the model components are different for the different regions Must make comparison on a regional basis. 

Southern North Sea has too few components (3) to do a meaningful PCA analysis, so PCA only carried out for Kattegat, Skagerrak and the Norwegian Trench  

###### Skagerrak PCA plot
```{r PCA Skagerrak, message=FALSE, warning=FALSE}
sc2$Sce_Mod <- paste(sc2$model, sc2$scenario, sep="  ")
#Skag_comb$Sce_Mod <- paste(Skag_comb$model, Skag_comb$scenario, sep="  ")

# to wide format
#am <- Skag_comb[,c(1,3,5)]
am <- sc2[,c(1,3,5)]
am$component<-as.factor(am$component)

tst <- am %>% 
  group_by(Sce_Mod, component) %>%
  summarise_all(funs(mean)) %>%
  pivot_wider(names_from = component, values_from= value, values_fill = 0)

rownames(tst) <- as.character(tst$Sce_Mod)
tst2 <- as.data.frame(t(tst[,c(2:17)]))
colnames(tst2) <- tst$Sce_Mod

 df_pca <- prcomp(tst2)
# df_pca2 <- prcomp(t(tst2))
#df_pca <- prcomp(tst[,c(2:17)])

# PCA biplot (using ggbiplot)
sk_pca <- ggbiplot(df_pca, labels =  rownames(t(tst2)),labels.size = 2.5, varname.size = 2.5, alpha=0.5) +coord_fixed(xlim=c(-3, 3), ylim=c(-3, 3)) + theme_bw() + ggtitle("Skagerrak")
sk_pca
ggsave("./figs/Skagerrak_PCA.png")


#### Loadings only plot
# Extract PC axes for plotting
PCAvalues <- data.frame(Scenarios = rownames(tst2), df_pca$x)

# Extract loadings of the variables
PCAloadings <- data.frame(Variables = rownames(df_pca$rotation), df_pca$rotation)

# proportion of variance
PoV <- df_pca$sdev^2/sum(df_pca$sdev^2)

# Plot
sk_load <- ggplot(PCAvalues, aes(x = PC1, y = PC2, colour = Species)) +
  geom_segment(data = PCAloadings, aes(x = 0, y = 0, xend = (PC1*5), yend = (PC2*5)), arrow = arrow(length = unit(1/2, "picas")), color = "gray60", linetype=11) +
  annotate("text", x = (PCAloadings$PC1*5), y = (PCAloadings$PC2*5), label = PCAloadings$Variables, color="darkblue") +
  coord_cartesian(xlim=c(-4, 4), ylim=c(-4, 1)) +
  theme_minimal() + 
  xlab(paste0("PC1 (", round((PoV[1]*100),digits=1), "%)")) +
  ylab(paste0("PC2 (", round((PoV[2]*100),digits=1), "%)")) +
  ggtitle("Skagerrak")
sk_load

ggsave("./figs/skagerrak_loadings.png")

```

##### Norwegian Trench PCA plot
```{r PCA NorwTrench, message=FALSE, warning=FALSE}
Nor_comb$Sce_Mod <- paste(Nor_comb$model, Nor_comb$scenario, sep="  ")

# to wide format
am <- Nor_comb[,c(1,3,5)]
am$component<-as.factor(am$component)

tst <- am %>% 
  group_by(Sce_Mod, component) %>%
  summarise_all(funs(mean)) %>%
  pivot_wider(names_from = component, values_from= value, values_fill = 0)

rownames(tst) <- as.character(tst$Sce_Mod)
tst2 <- as.data.frame(t(tst[,c(2:12)]))
colnames(tst2) <- tst$Sce_Mod

 df_pca <- prcomp(tst2)

# PCA biplot (using ggbiplot)
Norwt_pca <- ggbiplot(df_pca, labels =  rownames(tst2),labels.size = 2.5, varname.size = 2.5, alpha=0.5) +coord_fixed(xlim=c(-3, 3), ylim=c(-3, 3)) + theme_bw() +ggtitle("Norwegian Trench")
Norwt_pca
ggsave("./figs/NorwTrench_PCA.png")

#### Loadings only plot
# Extract PC axes for plotting
PCAvalues <- data.frame(Scenarios = rownames(tst2), df_pca$x)

# Extract loadings of the variables
PCAloadings <- data.frame(Variables = rownames(df_pca$rotation), df_pca$rotation)

# proportion of variance
PoV <- df_pca$sdev^2/sum(df_pca$sdev^2)

# Plot
nt_load <- ggplot(PCAvalues, aes(x = PC1, y = PC2, colour = Species)) +
  geom_segment(data = PCAloadings, aes(x = 0, y = 0, xend = (PC1*5), yend = (PC2*5)), arrow = arrow(length = unit(1/2, "picas")), color = "gray60", linetype=11) +
  annotate("text", x = (PCAloadings$PC1*5), y = (PCAloadings$PC2*5), label = PCAloadings$Variables, color="darkblue") +
  coord_cartesian(xlim=c(-4, 4.5), ylim=c(-4, 4)) +
  xlab(paste0("PC1 (", round((PoV[1]*100),digits=1), "%)")) +
  ylab(paste0("PC2 (", round((PoV[2]*100),digits=1), "%)")) +
  theme_minimal() +ggtitle("Norwegian Trench")
nt_load 
  
ggsave("./figs/NorwTrench_loadings.png")

```




##### Kattegat PCA plot
The seals X 10 dominates the PCA analysis, and is therefore excluded in the loadings plot to better illuminate how the other scenarios interrelate

```{r PCA Kattegat, message=FALSE, warning=FALSE}
Katt_comb$Sce_Mod <- paste(Katt_comb$model, Katt_comb$scenario, sep="  ")

# to wide format
am <- Katt_comb[,c(1,3,5)]
am$component<-as.factor(am$component)

tst <- am %>% 
  group_by(Sce_Mod, component) %>%
  summarise_all(funs(mean)) %>%
  pivot_wider(names_from = component, values_from= value, values_fill = 0)

rownames(tst) <- as.character(tst$Sce_Mod)
tst2 <- as.data.frame(t(tst[,c(2:14)]))
colnames(tst2) <- tst$Sce_Mod

 df_pca <- princomp(tst2, cor=TRUE)

# PCA biplot (using ggbiplot)
Katt_pca <- ggbiplot(df_pca, labels =  rownames(tst2),labels.size = 2.5, varname.size = 2.5, alpha=0.5, sec.axes = "cols") +coord_fixed(xlim=c(-3, 3), ylim=c(-3, 3)) + theme_bw() +ggtitle("Kattegat")
Katt_pca
ggsave("./figs/Kattegat_PCA.png")



df_pca %>%
  ggbiplot( sec.axes = "cols", scale.factor = 2, axis.percents=TRUE3) +
  theme_bw() +   expand_limits(y = c(-3.5, 3.5)) +   expand_limits(x = c(-3.5, 3.5))


#### Loadings only plot
# excluding the Seals X10 scenario

df_pca2 <- prcomp(tst2[c(1:4, 6:10)])
# Extract PC axes for plotting
PCAvalues <- data.frame(Scenarios = rownames(tst2), df_pca2$x)

# Extract loadings of the variables
PCAloadings <- data.frame(Variables = rownames(df_pca2$rotation), df_pca2$rotation)

# proportion of variance
PoV <- df_pca2$sdev^2/sum(df_pca2$sdev^2)

# Plot
ka_load <- ggplot(PCAvalues, aes(x = PC1, y = PC2, colour = Species)) +
  geom_segment(data = PCAloadings, aes(x = 0, y = 0, xend = (PC1*5), yend = (PC2*5)), arrow = arrow(length = unit(1/2, "picas")), color = "gray60", linetype=11) +
  annotate("text", x = (PCAloadings$PC1*5), y = (PCAloadings$PC2*5), label = PCAloadings$Variables, color="darkblue") +
  coord_cartesian(xlim=c(-3.7,3.5), ylim=c(-3, 2.5)) +
  theme_minimal() +
  xlab(paste0("PC1 (", round((PoV[1]*100),digits=1), "%)")) +
  ylab(paste0("PC2 (", round((PoV[2]*100),digits=1), "%)")) +
  ggtitle("Kattegat")
ka_load
  
ggsave("./figs/Kattegat_loadings.png")



```

#### Combine PCAs in one plot
```{r combine PCAs, echo=FALSE, warning= FALE}

#sk_pca + Katt_pca + Norwt_pca

load_comb <- sk_load / ka_load / nt_load
load_comb

ggsave("./figs/loadings_combined.png", unit="cm", width=20, height=40)

```

